---
title: "Aviation Cleaning"
author: "Lauren Puig"
format: pdf
editor: visual
---

# Adding Excel Sheets

```{r}
#| label: library
#| message: false


library(tidyverse)
library(readxl)
```

Steps to get data:

Went to: <https://data.ntsb.gov/avdata>

Acquired 9 most recent data advances for aircraft crashes. January 1, 2024 - March 15, 2024

Merged all together with excel files for the aircraft data and also the narratives

```{r}
#| label: adding-dataframes


aircraft_3 <- read_excel("ntsb/aircraft_3up01JAN.xlsx")
aircraft_2 <- read_excel("ntsb/aircraft_2.xlsx")
aircraft0 <- read_excel("ntsb/aircraft0.xlsx")
aircraft1 <- read_excel("ntsb/aircraft1.xlsx")
aircraft2 <- read_excel("ntsb/aircraft2.xlsx")
aircraft3 <- read_excel("ntsb/aircraft3.xlsx")
aircraft4 <- read_excel("ntsb/aircraft4.xlsx")
aircraft5 <- read_excel("ntsb/aircraft5.xlsx")
aircraft6 <- read_excel("ntsb/aircraft6.xlsx")

narratives_3 <- read_excel("ntsb/narratives_3up01JAN.xlsx")
narratives_2 <- read_excel("ntsb/narratives_2.xlsx")
narratives0 <- read_excel("ntsb/narratives0.xlsx")
narratives1 <- read_excel("ntsb/narratives1.xlsx")
narratives2 <- read_excel("ntsb/narratives2.xlsx")
narratives3 <- read_excel("ntsb/narratives3.xlsx")
narratives4 <- read_excel("ntsb/narratives4.xlsx")
narratives5 <- read_excel("ntsb/narratives5.xlsx")
narratives6 <- read_excel("ntsb/narratives6.xlsx")
```

Now need to merge the narratives with their ev_id. Not all accidents contain a narrative of what happen but usually over 50% do.

```{r}
#| label: merging-data

airnar_3 <- left_join(aircraft_3, narratives_3, by="ev_id")
airnar_2 <- left_join(aircraft_2, narratives_2, by="ev_id")
airnar0 <- left_join(aircraft0, narratives0, by="ev_id")
airnar1 <- left_join(aircraft1, narratives1, by="ev_id")
airnar2 <- left_join(aircraft2, narratives2, by="ev_id")
airnar3 <- left_join(aircraft3, narratives3, by="ev_id")
airnar4 <- left_join(aircraft4, narratives4, by="ev_id")
airnar5 <- left_join(aircraft5, narratives5, by="ev_id")
airnar6 <- left_join(aircraft6, narratives6, by="ev_id")
```

Duplicate in airnar5 - ev_id 20211221104432 has 2 instances because a plane collider with a paraglider and both had to be noted as one event.

Merge datasets into one!

```{r}
#| label: merge-one


MegaAirnar <- bind_rows(airnar_3, airnar_2, airnar0, airnar1, airnar2, airnar3, airnar4, airnar5, airnar6)

ValuableAirNar <- MegaAirnar |>
  select(ev_id, far_part, damage, acft_make, acft_model, acft_category, date_last_insp, owner_acft, oper_name, dprt_apt_id, dprt_city, dprt_state, dprt_country, dest_apt_id, dest_city, dest_state, dest_country, acft_year, fuel_on_board, narr_accp, narr_accf, narr_cause)

ValuableAirNar
```
